cardinalR: Generating interesting high-dimensional data structures

### 2025-09-19 paper-cardinalR.Rmd
Need to add a section for All noise in one
Updated mobius section with general noise dimensions
Updated pyramid one with noise dimensions

The Mobius strip is generated using the `geozoo::mobius` function with `p = 3`, and `n` defined as above. Each point is sampled from a parameterization defined by an angle $\theta \sim U(0, 2\pi)$, which specifies the position along the circular loop, and a width $w \sim U(-0.4, 0.4)$, which offsets the point across the strip’s band. The cartesian coordinates are $X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = (w/2)\sin(\theta/2)$. This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface. For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ Mobius into $p\text{-}D$, where $j = 4, \dots, p$.

@@

For all pyramid shapes, if $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1$. This tapering ensures that the structure narrows in all directions as it approaches the tip.

@@

>>>>

The Mobius strip is generated using the `geozoo::mobius` function with `p = 3`, and `n` defined as above. Each point is sampled from a parameterization defined by an angle $\theta \sim U(0, 2\pi)$, which specifies the position along the circular loop, and a width $w \sim U(-0.4, 0.4)$, which offsets the point across the strip’s band. The cartesian coordinates are $X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = (w/2)\sin(\theta/2)$. This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface. For $p > 3$, additional noise dimension are added to embed the $3\text{-}D$ Mobius into $p\text{-}D$.

@@

For all pyramid shapes, if $p > 3$, the remaining $p - 3$ dimensions (i.e., $X_4$ to $X_{p-1}$) are additional noise.

@@

### Generate noise dimensions

### 2025-09-19 paper-cardinalR.tex
Updated pyramid one with noise dimensions

For all pyramid shapes, if \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1\). This tapering ensures that the structure narrows in all directions as it approaches the tip.

>>>>

For all pyramid shapes, if \(p > 3\), the remaining \(p - 3\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are additional noise.
